% !TeX spellcheck = en_US
\section*{Analysis of memory usage}

Analysis report

by E. Marquer, 2018/05/23, Synalp and Université de Lorraine

\subsection{Abstract}

Experimental results shows (using \lstinline!nvidia-smi!) an increasing
memory usage for 4 layers, from an already enormous 3GB to more than
6GB, causing an out-of-memory error.

The objective of the following computations is to estimate the memory
consumption of the model, to confirm the hypothesis of a memory leak,
and verify that the model should not overflow memory.

\subsection{Formulas}

\subsubsection{Tensor and Variable size
estimation}

Byte size of a tensor is close to 6 times the products of all of its
dimensions. Byte size of a variable is similar to that of the
corresponding tensor.

\begin{lstlisting}[language=Python]
import torch, pickle

# Object to mesure
o = torch.autograd.Variable(torch.ones(100, 100, 100))
o = torch.ones(100, 100, 100)

len(pickle.dumps(o, 0)) / (100 * 100 * 100)
#result = 6
\end{lstlisting}

\subsubsection{Computations}
\vspace{-2em}
\begin{equation}
\begin{aligned}
total 
&= hidden\_states + msnn\_weights + emb\_weights + out\_weights \\\\
&= detach\_interval * (growth\_factor + 1) * layers * 6 * (hidden\_size * batch\_size * sequence\_length) \\
& + (((layer - 1) * 8 * hidden\_size * (hidden\_size + 1)) \\
& + 4 * hidden\_size * (hidden\_size + emb\_size + 2)) * 6  \\
& + (nwords * (emb\_size + 1)) * 6 \\
& + (nwords * (layers * hidden\_size)) * 6 \\\\
&= 6 * (detach\_interval * (growth\_factor + 1) * layers * 
(hidden\_size * batch\_size * sequence\_length) \\
& + ((layers - 1) * 8 * hidden\_size * (hidden\_size + 1)) \\
& + 4 * hidden\_size * (hidden\_size + emb\_size + 2)  \\
& + (nwords * (emb\_size + 1)) \\
& + (nwords * (layers * hidden\_size)))
\end{aligned}
\end{equation}

\begin{equation}
\begin{aligned}
hidden\_states &= total\_history * 6 * dim \\
               &= detach\_interval * (growth\_factor + 1) * layers * 6 * dim \\
               &= detach\_interval * (growth\_factor + 1) * layers * 6 * \\
               & (hidden\_size * batch\_size * sequence\_length)
\end{aligned}
\end{equation}

\begin{equation}
\begin{aligned}
msnn\_layer\_weights &= weights\_ih + weights\_hh + bias\_ih + bias\_hh \\
                     &= 4 * hidden\_size * input\_size + 4 * hidden\_size * hidden\_size \\
                     & + 4 * hidden\_size + 4 * hidden\_size \\
                     &= 4 * hidden\_size * (hidden\_size + input\_size + 2) \\
                     &= \begin{cases}
                        8 * hidden\_size * (hidden\_size + 1) &\text{for all layers exept the first one} \\
                        4 * hidden\_size * (hidden\_size + emb\_size + 2)  &\text{for the first layer}
                        \end{cases}
\end{aligned}
\end{equation}

\begin{equation}
\begin{aligned}
msnn\_weights &= msnn\_layer\_weights * layers \\
              &= ((layers - 1) * 8 * hidden\_size * (hidden\_size + 1))\\
              & + 4 * hidden\_size * (hidden\_size + emb\_size + 2)
\end{aligned}
\end{equation}

\begin{equation}
\begin{aligned}
emb\_weights &= (bias + weights) * 6 \\
            &= (nwords * emb\_size + emb\_size))* 6 \\
            &= (nwords * (emb\_size + 1)) * 6
\end{aligned}
\end{equation}

\begin{equation}
out\_weights = (nwords * (layers * hidden\_size)) * 6
\end{equation}

\newpage
\subsubsection{Estimate with basic set of parameters}

\begin{lstlisting}[language=Python]
detach_interval = 50
growth_factor = 5
layers = 7
hidden_size = 1840 / 4
batch_size = 2
sequence_length = 100
emb_size = 400
nwords = 205

total = 6 * (detach_interval * (growth_factor + 1) * layers * hidden_size * batch_size * sequence_length + ((layers - 1) * 8 * hidden_size * (hidden_size + 1)) + 4 * hidden_size * (hidden_size + emb_size + 2) + (nwords * (emb_size + 1)) + (nwords * (layers * hidden_size)))

"{}GB {}MB {}kB {}B".format(int(total%(1024**4) / 1024**3), int(total%(1024**3) / 1024**2), int(total%(1024**2) / 1024), int(total%1024))
# result: '1GB 741MB 614kB 9B'
\end{lstlisting}

\newpage
\rotatebox{90}{
\begin{tabular}{|cccccccc|r|}
\hline
detach\_interval & growth\_factor & layers & hidden\_size & batch\_size & sequence\_length & emb\_size & nwords & total\\
\hline
50 & 5 & 7 & 1840 / 4 & 2 & 200 / batch\_size & 400 & 205 & 1GB 153MB 68kB 6B\\
\textbf{100} & 5 & 7 & 1840 / 4 & 2 & 200 / batch\_size & 400 & 205 & 2GB 234MB 579kB 262B\\
\textbf{200} & 5 & 7 & 1840 / 4 & 2 & 200 / batch\_size & 400 & 205 & 4GB 397MB 577kB 774B\\
50 & \textbf{10} & 7 & 1840 / 4 & 2 & 200 / batch\_size & 400 & 205 & 2GB 50MB 323kB 390B\\
50 & 5 & \textbf{6} & 1840 / 4 & 2 & 200 / batch\_size & 400 & 205 & 1008MB 912kB 414B\\
50 & 5 & \textbf{5} & 1840 / 4 & 2 & 200 / batch\_size & 400 & 205 & 840MB 732kB 822B\\
50 & 5 & \textbf{4} & 1840 / 4 & 2 & 200 / batch\_size & 400 & 205 & 672MB 553kB 206B\\
50 & 5 & \textbf{3} & 1840 / 4 & 2 & 200 / batch\_size & 400 & 205 & 504MB 373kB 614B\\
50 & 5 & \textbf{2} & 1840 / 4 & 2 & 200 / batch\_size & 400 & 205 & 336MB 193kB 1022B\\
50 & 5 & \textbf{1} & 1840 / 4 & 2 & 200 / batch\_size & 400 & 205 & 168MB 14kB 406B\\
50 & 5 & 7 & \textbf{1840 / 2} & 2 & 200 / batch\_size & 400 & 205 & 2GB 431MB 598kB 94B\\
50 & 5 & 7 & \textbf{1840 / 8} & 2 & 200 / batch\_size & 400 & 205 & 573MB 28kB 890B\\
50 & 5 & 7 & 1840 / 4 & \textbf{1} & \textbf{200 / batch\_size} & 400 & 205 & 1GB 153MB 68kB 6B\\
50 & 5 & 7 & 1840 / 4 & 2 & \textbf{200} & 400 & 205 & 2GB 234MB 579kB 262B\\
50 & 5 & 7 & 1840 / 4 & 2 & 200 / batch\_size & \textbf{200} & 205 & 1GB 150MB 743kB 534B\\
50 & 5 & 7 & 1840 / 4 & 2 & 200 / batch\_size & \textbf{800} & 205 & 1GB 157MB 764kB 998B\\
50 & 5 & 7 & 1840 / 4 & 2 & 200 / batch\_size & 400 & \textbf{500} & 1GB 159MB 182kB 984B\\
\hline
\end{tabular}
}

\newpage
\paragraph{Most impactful factors
(memory-wise)}

\begin{itemize}
\item
  \lstinline!detach_interval!: \lstinline!detach_interval * 2! =
  \lstinline!memory * 2!
\item
  \lstinline!growth_factor!: \lstinline!growth_factor * 2! =
  \lstinline!memory * 2!
\item
  \lstinline!hidden_size!: \lstinline!hidden_size * 2! =
  \lstinline!memory * 2!
\item
  \lstinline!batch_size * sequence_length! (number of examples by
  sequence): \lstinline!batch_size*sequence_length * 2! =
  \lstinline!memory * 2!
\item
  \lstinline!layers!: \lstinline!layers * 2! = \lstinline!layers * 2!
\end{itemize}

The others factors considered (emb\_size and nwords) have almost no
impact on memory. It confirms that the most memoryphage element is the
MSNN. Also, to keep a stable memory usage,
\lstinline!batch_size * sequence_length! ratio must be kept constant;
increasing \lstinline!batch_size! while lowering
\lstinline!sequence_length! can increase processing speed whithout
impacting memory usage.

\subparagraph{Impact of layer increase}
\subparagraph*{}\vspace{-2em}{We can notice that the impacts of layers is most noticeable during the
first phases of training, during the creation of the first layer. Later
on, the creation frequency of new layers is small, and the change is
minimal. For example, from 6 to 7 layer, we need \(`12500`\) sequences
to pass, and the increase of memory of about \(`7/6`\); from 7 to 8
layer, we need \(`62500`\) sequences to pass, and the increase of memory
of about \(`8/7`\); from 8 to 9 layer, we need \(`312500`\) sequences to
pass, and the increase of memory of about \(`9/8`\); and so on.}

\subsubsection{Partial derivatives:}

\begin{lstlisting}
d = detach_interval
g = growth_factor
l = layers
h = hidden_size
b = batch_size
s = sequence_length
e = emb_size
n = nwords

complete formula:
6 * (
	d * (g + 1) * l * h * b * s +
	((l - 1) * 8 * h * (h + 1)) +
	4 * h * (h + e + 2) +
	(n * (e + 1)) +
	(n * (l * h))
)

simplified formula:
6*(8*l*h^2 + l*d*g*h*b*s + l*d*h*b*s + 8*l*h + l*n*h + 4*e*h + e*n + n - 4*h^2)

partial derivates on given variable:
d: (6(8*l*h^2 - 4*h^2 + l*d*g*h*b*s + l*d*h*b*s + 8*l*h + 4*e*h + l*n*h + e*n + n))
g: 6*l*d*h*b*s
l: 6*(d*h*b*s*(g+1) + 8*h*(h+1) + nh)
h: 6*(l*d*g*b*s + l*d*b*s + l*n + 16*l*h + 8*l + 4*e - 8*h)
b: 6*l*d*h*s*(g+1)
s: 6*l*d*h*b*(g+1)
e: 0
n: 6*(l*h+e+1)
\end{lstlisting}

